Local real space operators for translating Stokes Q/U to E/B maps on the sphere

### 2017-10-17 article/sections/kernel_discussion.tex
updates to discussion on locality of E and B mode kernels

\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
It is clear from previous discussions that that the non-locality of the E and B modes is determined by the radial part of the convolution kernels. To quantify this non-locality as a function of the maximum multipole accessible for analysis, we evaluate the radial part of the convolution kernel for different values of $\ell_{\rm max}$, while keeping the lowest multipole fixed at $\ell_{\rm min}=2$. The set of radial kernels so derived are plotted in \fig{fig:rad_ker_decay}. All the function have been normalized such that their global maxima is set to unity. Note that on increasing $\ell_{\rm max}$ the radial kernels shift to the left, attaining fractions of  their global maxima at progressively small angular distances $\beta$ from the central pixel. 
%
\begin{figure}[!t]
\centering
\subfigure[]{\includegraphics[width=1.0\columnwidth]{kernel/f_rad_ker_fn_of_ellmax.pdf}}
\subfigure[]{\includegraphics[width=0.48\columnwidth]{kernel/fp2_rad_ker_fn_of_ellmax.pdf}}
\subfigure[]{\includegraphics[width=0.48\columnwidth]{kernel/fm2_rad_ker_fn_of_ellmax.pdf}}
\caption{The top panel shows a plot of the radial kernels $f(\beta,\ell_{\rm min},\ell_{\rm max})$ while the bottom left and right panels show the radial functions ${}_{+2}f(\beta,\ell_{\rm min},\ell_{\rm max}) ~\&~ {}_{-2}f(\beta,\ell_{\rm min},\ell_{\rm max})$ respectively, for different $\ell_{\rm max}$ as indicated by the legends and fixed $\ell_{\rm min}=2$. The curves for each of these functions have been normalized such that the maximum of the curve is set to unity. The horizontal solid black line marks the location where the amplitude of the kernel falls below 1\% of its maximum. The slanted dashed black lines indicate a power law fit (by eye) to the envelope of the radial functions. While the envelopes for function $f(\beta)~\&~ {}_{-2}f(\beta)$ are fit well by the power law $\propto \beta^{-1.5}$, the envelope for the function ${}_{-2}f(\beta)$ is seen to have a slightly steeper fall off $\propto \beta^{-1.65}$.}
\label{fig:rad_ker_decay}
\end{figure}
%
We define the value of the abscissa at which the function $f(\beta,\ell_{\rm min},\ell_{\rm max})$ transits to being monotonously below 1\% of the maxima of the function as the non-locality parameter: $\beta_{o}$. We find that the following empirical relation: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$ can predict quite accurately the value of the non-locality parameter for a given maximum multipole $\ell_{\rm max}$ and fixed $\ell_{\rm min}=2$.

The envelope of the radial kernel $f(\beta,\ell_{\rm min},\ell_{\rm max})$ is observed to have a linear relation to the angular distance $\beta$ in log space as seen in \fig{fig:rad_ker_decay}, indicating a power law fall off. A fit by eye indicates that the envelope of the function is well represented by a fuction proportional to $\beta^{-1.5}$ in intermediate angular distance range. This is not valid in regions $\beta << \beta_o$ and $\beta >> \beta_o$, where the envelope of the function can be clearly seen to deviate from power law behavior (see \fig{fig:rad_ker_decay}).

\comment{There isn't too much of a discussion surrounding the function ${}_{\pm}f(\beta)$.}
\comment{Comment on the telescoping behavior of these radial functions ? }

>>>>

\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
It is clear from previous discussions that that the non-locality of the E and B modes is determined by the radial part of the convolution kernels. To quantify this non-locality as a function of the maximum multipole accessible for analysis, we evaluate the radial part of the convolution kernel for different values of $\ell_{\rm max}$, while keeping the lowest multipole fixed at $\ell_{\rm min}=2$. 
%
\begin{figure}[t]
\centering
\subfigure{\includegraphics[width=0.96\columnwidth]{kernel/f_rad_ker_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{kernel/fp2_rad_ker_fn_of_ellmax.pdf}}
\subfigure{\includegraphics[width=0.48\columnwidth]{kernel/fm2_rad_ker_fn_of_ellmax.pdf}}
\caption{The top panel shows a plot of the radial kernels $f(\beta,\ell_{\rm min},\ell_{\rm max})$ while the bottom left and right panels show the radial functions ${}_{+2}f(\beta,\ell_{\rm min},\ell_{\rm max}) ~\&~ {}_{-2}f(\beta,\ell_{\rm min},\ell_{\rm max})$ respectively, for different $\ell_{\rm max}$ as indicated by the legends and fixed $\ell_{\rm min}=2$. The curves for each of these functions have been normalized such that the maximum of the curve is set to unity. The horizontal solid black line marks the location where the amplitude of the kernel falls below 1\% of its maximum. The slanted dashed black lines indicate a power law fit (by eye) to the envelope of the radial functions. While the envelopes for function $f(\beta)~\&~ {}_{-2}f(\beta)$ are fit well by the power law $\propto \beta^{-1.5}$, the envelope for the function ${}_{-2}f(\beta)$ is seen to have a slightly steeper fall off $\propto \beta^{-1.65}$.}
\label{fig:rad_ker_decay}
\end{figure}
%
The set of radial kernels so derived are plotted in \fig{fig:rad_ker_decay}. All the function have been normalized such that their global maxima is set to unity. Note that on increasing $\ell_{\rm max}$ the radial kernels shift to the left, attaining fractions of  their global maxima at progressively small angular distances $\beta$ from the central pixel.

We have observed that the radial functions derived by evaluating the multipole sums to different maximum multipoles are self similar. Specifically these functions follow an interesting telescoping and scaling property: $${}_rf(\beta,2,\ell_{\rm max}) \approx [\frac{\ell_{\rm max}}{\ell'_{\rm max}}]^2{}_rf(\beta'=\frac{\ell_{\rm max}}{\ell'_{\rm max}} \beta ,2,\ell'_{\rm max}) \,,$$ where ${}_rf$ denotes all the different radial functions. We can understand the shifting left of the radial kernels on increasing the maximum multipole using this property. Lets say the function ${}_rf(\beta, \ell'_{\rm max})$ transition to being monotonously below some fraction of the global maxima at an angular distance of $\beta_0$.  The function ${}_rf(\beta', \ell_{\rm max})$, given $\ell_{\rm max} > \ell'_{\rm max}$, reaches the same transition point $\beta'=\beta_0$ at a smaller value of $\beta$ owing to the fact that $\ell_{\rm max}/\ell'_{\rm max}$ is greater than unity. The amplitude scaling of the functions is irrelevant since the transition point is always described in terms of the fraction of the global maxima of the function.

We are primarily interested in the non-local dependence of the scalar modes E \& B on the Stokes parameters Q \& U. We define the value of the abscissa at which the function $f(\beta,\ell_{\rm min},\ell_{\rm max})$ transits to being monotonously below 1\% of the maxima of the function as the non-locality parameter: $\beta_{o}$.  For a $\ell_{\rm max}=24$, the maximum multipole accessible on a Nside=8 Healpix map, the non-locality parameter $\beta_0=180^{\circ}$, since the radial function never falls monotonously below 1\% of the global maxima of the function. Using the self similar  property of the radial function discussed above, we define the following: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$ as the non-locality parameter, given the maximum multipole $\ell_{\rm max}$. \revisit{$\beta_0$ is not very sensitive to $\ell_{\rm min}$, though it can be important on making drastic changes to $\ell_{\rm min}$. } \comment{Quantify how changing $\ell_{\rm min}$ changes $\beta_0$}

The envelope of the radial kernel $f(\beta,\ell_{\rm min},\ell_{\rm max})$ is observed to have a power law relation to the angular distance $\beta$ as seen in \fig{fig:rad_ker_decay}. At small values of $\beta$, which corresponds to the flat sky limit,  the radial function is proportional to $\beta^2$. In the intermediate range,  a fit by eye indicates that the envelope of the function is well represented by the power law $\beta^{-1.5}$.
\comment{There isn't too much of a discussion surrounding the function ${}_{\pm}f(\beta)$.}
